
This file converts Joop's comments_mods.do to R with the help of ChatGPT 4o.

# Dependencies and data

```{r}

# dependencies
library(tidyverse)
library(plm)
library(broom)
library(haven)
library(parameters)
library(ggplot2)
library(ggstance)
library(lmtest)

# load data
#dat1 <- read_dta("../0 original materials/DRMCoronaWave1DTA_EPISODE_CLEAN.dta") # used in Joop's code, but not present in OSF materials
#dat2 <- read_dta("../0 original materials/DRMCoronaWave1DTA_BJHP.dta") # present in OSF materials
dat3 <- read_dta("../0 original materials/DRMCoronaWave1DTA_EPISODE_BJHP.dta") # present in OSF materials. seems to contain the variables needed for the analysis.
#dat4 <- read_dta("../0 original materials/DRMCoronaWave1DTA_CLEAN.dta") # present in OSF materials

# use dat3
data <- dat3

```

# Part I: Standard Error Corrections

Joop's comment: "For panel (a) of Figure I,  this change renders 1 point estimate insignificant at the 5% level for the positive affect outcome."

```{r fig.height=8, fig.width=6}

# Convert to panel data
pdata <- pdata.frame(data, index = "id")

# Run fixed effects regressions
model1 <- plm(zposaff ~ activity_1E + activity_2E + activity_3E + activity_4E + activity_5E + 
                activity_6E + activity_7E + activity_8E + activity_9E + activity_10E + 
                activity_11E + activity_12E + activity_13E + activity_14E + activity_15E + 
                activity_16E + activity_17E + activity_18E + activity_19E + activity_20E + 
                activity_21E + activity_22E + activity_23E + activity_24E + activity_25E, 
              data = pdata, model = "within")

model2 <- plm(zposaff ~ activity_1E + activity_2E + activity_3E + activity_4E + activity_5E + 
                activity_6E + activity_7E + activity_8E + activity_9E + activity_10E + 
                activity_11E + activity_12E + activity_13E + activity_14E + activity_15E + 
                activity_16E + activity_17E + activity_18E + activity_19E + activity_20E + 
                activity_21E + activity_22E + activity_23E + activity_24E + activity_25E, 
              data = pdata, model = "within", vcov = vcovHC, cluster = "id")

# Summary of the models
model_parameters(model1)
model_parameters(model2)

# compare results
dat_combined_zposaff <- bind_rows(
  model_parameters(model1) |>
    mutate(model = "original"),
  model_parameters(model2) |>
    mutate(model = "modified")
)

ggplot(dat_combined_zposaff, aes(Coefficient, Parameter, color = model)) +
  geom_vline(xintercept = 0) +
  geom_linerangeh(aes(xmin = CI_low, xmax = CI_high), position = position_dodge(width = .5)) +
  geom_point(position = position_dodge(width = .5)) +
  scale_color_viridis_d(begin = 0.3, end = 0.7) +
  ggtitle("Positive affect - Standard Error Corrections")

```

```{r fig.height=8, fig.width=6}

model3 <- plm(znegaff ~ activity_1E + activity_2E + activity_3E + activity_4E + activity_5E + 
                activity_6E + activity_7E + activity_8E + activity_9E + activity_10E + 
                activity_11E + activity_12E + activity_13E + activity_14E + activity_15E + 
                activity_16E + activity_17E + activity_18E + activity_19E + activity_20E + 
                activity_21E + activity_22E + activity_23E + activity_24E + activity_25E, 
              data = pdata, model = "within")

model4 <- plm(znegaff ~ activity_1E + activity_2E + activity_3E + activity_4E + activity_5E + 
                activity_6E + activity_7E + activity_8E + activity_9E + activity_10E + 
                activity_11E + activity_12E + activity_13E + activity_14E + activity_15E + 
                activity_16E + activity_17E + activity_18E + activity_19E + activity_20E + 
                activity_21E + activity_22E + activity_23E + activity_24E + activity_25E, 
              data = pdata, model = "within", vcov = vcovHC, cluster = "id")

model_parameters(model3)
model_parameters(model4)

# compare results
dat_combined_znegaff <- bind_rows(
  model_parameters(model3) |>
    mutate(model = "original"),
  model_parameters(model4) |>
    mutate(model = "modified")
)

ggplot(dat_combined_znegaff, aes(Coefficient, Parameter, color = model)) +
  geom_vline(xintercept = 0) +
  geom_linerangeh(aes(xmin = CI_low, xmax = CI_high), position = position_dodge(width = .5)) +
  geom_point(position = position_dodge(width = .5)) +
  scale_color_viridis_d(begin = 0.3, end = 0.7) +
  ggtitle("Negative affect - Standard Error Corrections")

```

# Part II: Interpretation of "Reference" Category

```{r fig.height=9, fig.width=6}

pdata_modified <- pdata |>
  rowwise() |>
  mutate(N_act_ep = as.numeric(sum(c_across(starts_with("activity_")))),
         N_act_ep = as.factor(N_act_ep)) |>
  ungroup()

# Run regressions and store results
model5 <- plm(zposaff ~ activity_1E + activity_2E + activity_3E + activity_4E + activity_5E + 
                activity_6E + activity_7E + activity_8E + activity_9E + activity_10E + 
                activity_11E + activity_12E + activity_13E + activity_14E + activity_15E + 
                activity_16E + activity_17E + activity_18E + activity_19E + activity_20E + 
                activity_21E + activity_22E + activity_23E + activity_24E + activity_25E, 
              data = pdata, model = "within", vcov = vcovHC, cluster = "id")

# Choose resting/relaxing as reference category
model7 <- plm(zposaff ~ N_act_ep + activity_1E + activity_2E + activity_3E + activity_4E + activity_5E + 
                activity_6E + activity_7E + activity_8E + activity_9E + activity_10E + 
                activity_11E + activity_12E + activity_13E + activity_14E + activity_15E + 
                activity_16E + activity_17E + activity_18E + activity_19E + activity_20E + 
                activity_21E + activity_22E + activity_23E + activity_24E + activity_25E, 
              #data = pdata_modified, model = "within", vcov = vcovHC, cluster = "id") # throws error for some reason. vcovHC solution below suggested by chatGPT.
              data = pdata_modified, model = "within", effect = "individual")

# Adjust standard errors for clustering
model7_vcov <- vcovHC(model7, method = "arellano", type = "HC0", cluster = "group")

# compare results
dat_combined_zposaff <- bind_rows(
  model_parameters(model5) |>
    mutate(model = "original"),
  model_parameters(coeftest(model7, model7_vcov)) |>
    mutate(model = "modified")
)

ggplot(dat_combined_zposaff, aes(Coefficient, Parameter, color = model)) +
  geom_vline(xintercept = 0) +
  geom_linerangeh(aes(xmin = CI_low, xmax = CI_high), position = position_dodge(width = .75)) +
  geom_point(position = position_dodge(width = .75)) +
  scale_color_viridis_d(begin = 0.3, end = 0.7) +
  ggtitle("Positive affect - Interpretation of Reference Category")

```

```{r fig.height=9, fig.width=6}

model6 <- plm(znegaff ~ activity_1E + activity_2E + activity_3E + activity_4E + activity_5E + 
                activity_6E + activity_7E + activity_8E + activity_9E + activity_10E + 
                activity_11E + activity_12E + activity_13E + activity_14E + activity_15E + 
                activity_16E + activity_17E + activity_18E + activity_19E + activity_20E + 
                activity_21E + activity_22E + activity_23E + activity_24E + activity_25E, 
              data = pdata, model = "within", vcov = vcovHC, cluster = "id")

model8 <- plm(znegaff ~ N_act_ep + activity_1E + activity_2E + activity_3E + activity_4E + activity_5E + 
                activity_6E + activity_7E + activity_8E + activity_9E + activity_10E + 
                activity_11E + activity_12E + activity_13E + activity_14E + activity_15E + 
                activity_16E + activity_17E + activity_18E + activity_19E + activity_20E + 
                activity_21E + activity_22E + activity_23E + activity_24E + activity_25E, 
              #data = pdata_modified, model = "within", vcov = vcovHC, cluster = "id") # throws error for some reason. vcovHC solution below suggested by chatGPT.
              data = pdata_modified, model = "within", effect = "individual")

model8_vcov <- vcovHC(model8, method = "arellano", type = "HC0", cluster = "group")

# compare results
dat_combined_znegaff <- bind_rows(
  model_parameters(model6) |>
    mutate(model = "original"),
  #model_parameters(model8)
  model_parameters(coeftest(model8, model8_vcov)) |>
    mutate(model = "modified")
)

ggplot(dat_combined_znegaff, aes(Coefficient, Parameter, color = model)) +
  geom_vline(xintercept = 0) +
  geom_linerangeh(aes(xmin = CI_low, xmax = CI_high), position = position_dodge(width = .75)) +
  geom_point(position = position_dodge(width = .75)) +
  scale_color_viridis_d(begin = 0.3, end = 0.7) +
  ggtitle("Negative affect - Interpretation of Reference Category")

```

- why do the models label activity_1E differently when they're specified the same?
- why is activity_25E missing for for the modified model despite being specified?

# Part III: Absence of "Persons"

NEEDS WORK. whats the comparison?

```{r fig.height=6, fig.width=6}

# Generating variables for presence/absence of persons
pdata <- pdata %>% 
  mutate(total_perswith = spousePers + KidsPers + FriendsPers + ParentsPers + WorkPers + PetsPers + OtherPers + NobodyPers,
         total_distwith = spouseDist + KidsDist + FriendsDist + ParentsDist + WorkDist + PetsDist + OtherDist + NobodyDist,
         someone_perswith = (spousePers + KidsPers + FriendsPers + ParentsPers + WorkPers + PetsPers + OtherPers) > 0,
         someone_distwith = (spouseDist + KidsDist + FriendsDist + ParentsDist + WorkDist + PetsDist + OtherDist) > 0,
         Nobody = (spousePers == 0 & spouseDist == 0 & KidsPers == 0 & KidsDist == 0 & FriendsPers == 0 & FriendsDist == 0 & ParentsPers == 0 & ParentsDist == 0 & WorkPers == 0 & WorkDist == 0 & PetsPers == 0 & PetsDist == 0 & OtherPers == 0 & OtherDist == 0))

# Run regressions
model9 <- plm(zposaff ~ spousePers + spouseDist + KidsPers + KidsDist + FriendsPers + FriendsDist + 
                ParentsPers + ParentsDist + WorkPers + WorkDist + PetsPers + PetsDist + OtherPers + OtherDist + NobodyPers, 
              data = pdata, model = "within")

# Summary of the models
model_parameters(model9)

# # compare results
# dat_combined_zposaff <- bind_rows(
#   model_parameters(model9) |>
#     mutate(model = "original"),
#   model_parameters(model10) |>
#     mutate(model = "modified")
# )
# 
# ggplot(dat_combined_zposaff, aes(Coefficient, Parameter, color = model)) +
#   geom_vline(xintercept = 0) +
#   geom_linerangeh(aes(xmin = CI_low, xmax = CI_high), position = position_dodge(width = .5)) +
#   geom_point(position = position_dodge(width = .5)) +
#   scale_color_viridis_d(begin = 0.3, end = 0.7)

```

```{r fig.height=6, fig.width=6}

# Run regressions
model10 <- plm(znegaff ~ spousePers + spouseDist + KidsPers + KidsDist + FriendsPers + FriendsDist + 
                 ParentsPers + ParentsDist + WorkPers + WorkDist + PetsPers + PetsDist + OtherPers + OtherDist + NobodyPers, 
               data = pdata, model = "within")

# Summary of the models
model_parameters(model10)

# # compare results
# dat_combined_znegaff <- bind_rows(
#   model_parameters(model9) |>
#     mutate(model = "original"),
#   model_parameters(model10) |>
#     mutate(model = "modified")
# )
# 
# ggplot(dat_combined_znegaff, aes(Coefficient, Parameter, color = model)) +
#   geom_vline(xintercept = 0) +
#   geom_linerangeh(aes(xmin = CI_low, xmax = CI_high), position = position_dodge(width = .5)) +
#   geom_point(position = position_dodge(width = .5)) +
#   scale_color_viridis_d(begin = 0.3, end = 0.7)

```

# BH corrections on p values

```{r}

model_parameters(model8) |> # is this the model to adjust?
  as_tibble() |>
  mutate(p_adj = p.adjust(p, method = "BH"),
         p_sig = p < .05,
         p_adj_sig = p_adj < .05) |>
  mutate_if(is.numeric, round_half_up, digits = 2)

```

